Governing through patronage: the bargain for education in decentralized Brazil

### 2020-11-16 paper/paper.tex
fix principal descriptives labels

In Brazil, the responsibility to provide public, primary education is primarily delegated to local governments \citet{paschoal_historia_2009}. In this critical learning period, students acquire skills such as reading/writing, as well as foundational concepts in math such as addition and substraction. The municipal educational system has increased in relevance over the past few decades, and currently over 60 percent of lower school strudents attend a public school. Figure \ref{fig:enroll} plots the total number of students in primary education per government level, including private schools. As of 2016, over 25 million students were enrolled in over 115 thousand municipal schools.

@@

Teachers and school principals are responsible for providing all educational services at the local level. As of 2016, in the municipal education system there were 3.7 million teachers distributed across the national territory, as well as 148 thousand school principals. TableA majority of teachers are hired with a permanent contract (\emph{estatutário}) Over the past decade, there has been an increasing

>>>>

In Brazil, the responsibility to provide public, primary education is primarily delegated to local governments \citet{paschoal_historia_2009}. In this critical learning period, students acquire skills such as reading/writing, as well as foundational concepts in mathematics. The municipal educational system has increased in size over the past few decades, and currently over 60 percent of lower school students attend a public school. Figure \ref{fig:enroll} plots the total number of students in primary education per government level, including private schools. As of 2016, over 25 million students were enrolled in over 115 thousand municipal schools.

@@

Teachers and school principals are responsible for providing all educational services at the local level, overseen by a local department of education (\emph{secretaria de educa\c{c}\~{a}o}). As of 2016, in the municipal education system there were 3.7 million teachers distributed across the national territory, as well as 148 thousand school principals. A majority of teachers are hired with a permanent contract (\emph{estatutário}), with over 73 percent of teachers enrolled with civil service protection. This contrasts with the hiring of school principals, who are politically appointed and therefore have no similar protections set in place.

\begin{table}[t]
    \input{results/descriptive_teacher}
    \caption{\textbf{Summary statistics for municipal teachers.} All statistics take as reference year 2017.}
    \label{tbl:descriptive_teacher}
\end{table}

Table \ref{tbl:descriptive_teacher} provides some additional descriptive statistics regarding teachers. The teacher labor force is highly educated and primarily composed of women, and schools in Brazil are relatively small, with an average of 4 teachers per school. As for school principals, while

\begin{figure}[h]
    \centering
    \includegraphics{plots/plot_principal_descriptive.pdf}
\end{figure}

### 2020-11-16 paper/results/descriptive_teacher.tex
fix principal descriptives labels

% Table created by stargazer v.5.2.2 by Marek Hlavac, Harvard University. E-mail: hlavac at fas.harvard.edu
% Date and time: Mon, Nov 16, 2020 - 10:49:28 AM
\begin{tabular}{@{\extracolsep{5pt}}lcccc} 
\\[-1.8ex]\hline 
\hline \\[-1.8ex] 
Statistic & \multicolumn{1}{c}{Mean} & \multicolumn{1}{c}{St. Dev.} & \multicolumn{1}{c}{Min} & \multicolumn{1}{c}{Max} \\ 
\hline \\[-1.8ex] 
Number of teachers per school & 4.108 & 4.335 & 1 & 86 \\ 
Number of classes per school & 4.126 & 4.417 & 1 & 173 \\ 
Proportion of teachers w/ higher education & 0.729 & 0.377 & 0 & 1 \\ 
Average age & 0.811 & 0.283 & 0.000 & 1.000 \\ 
Average number of classes per teacher & 40.008 & 7.251 & 16.000 & 75.000 \\ 
mean\_n\_class\_per\_teacher & 1.025 & 0.248 & 0 & 7 \\ 
\hline \\[-1.8ex] 
\end{tabular}

>>>>

% Table created by stargazer v.5.2.2 by Marek Hlavac, Harvard University. E-mail: hlavac at fas.harvard.edu
% Date and time: Mon, Nov 16, 2020 - 11:05:29 AM
\begin{tabular}{@{\extracolsep{5pt}}lcccc} 
\\[-1.8ex]\hline 
\hline \\[-1.8ex] 
Statistic & \multicolumn{1}{c}{Mean} & \multicolumn{1}{c}{St. Dev.} & \multicolumn{1}{c}{Min} & \multicolumn{1}{c}{Max} \\ 
\hline \\[-1.8ex] 
Number of teachers per school & 4.108 & 4.335 & 1 & 86 \\ 
Number of classes per school & 4.126 & 4.417 & 1 & 173 \\ 
Proportion of teachers w/ higher education & 0.729 & 0.377 & 0 & 1 \\ 
Average age & 40.008 & 7.251 & 16.000 & 75.000 \\ 
Proportion of women & 0.811 & 0.283 & 0.000 & 1.000 \\ 
Average number of classes per teacher & 1.025 & 0.248 & 0 & 7 \\ 
\hline \\[-1.8ex] 
\end{tabular}
